---
title: "Assignment3"
author: "Jiaqiu Wang"
date: "March 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part I: OLS Regression

```{r}
rm(list = ls())
setwd("C:/Users/rache/Google Drive/Spring 2017 Academic/fintech/assignment/ass3")
library("glmnet")
library("dplyr")
Data_1970 <- read.csv("Data_Daily_1970.csv",header=TRUE) 
head(Data_1970)
```

```{r}
industry_return <- paste("FFInd",1:49, sep="")
lag_industy_return <- apply(Data_1970[,industry_return], 2, lag)
lag_industy_return <- lag_industy_return[-1,]

mkt_excess <-Data_1970$mkt_excess
mkt_excess <- mkt_excess[-1]

model1 <- lm(mkt_excess ~ lag_industy_return)
summary(model1)
ols_coef <- coef(model1)
```
R-squared is 0.3564, over-all fitting p-value is 2.899e-05, it appears that the lagged portfolio
returns can predict the excess return of US equity market.

## Part II: Lasso Regression

```{r}
model2 <- cv.glmnet(lag_industy_return,mkt_excess,alpha=1,standardize=FALSE)
bestlam <- model2$lambda.min
lasso_coef <- coef(model2, s = bestlam)
lasso_coef <- as.numeric(lasso_coef)
```

```{r}
#plot histgram for both regression coefficients
hist(ols_coef,plot=TRUE)
hist(lasso_coef,plot = TRUE)
```

From the histgram we can see that compared with OLS coefficients, the coefficient distribution from Lasso regression has more zeros, and the non-zero coefficients tend to be further away from zero.


## Part 3: Out-of-sample trading experiment
With the OLS regression coefficients, predict the daily returns in 1971 using lag data.Then construct the portfolio using the simple strategy: long the US equity market if the predicted return is positive, or short if not. Calculate the gross excess return of your portfolio, as well as the Sharpe Ratio using daily returns.

```{r}
Data_1971 <- read.csv("Data_Daily_1971.csv",header=TRUE) 
#head(Data_1971)
lag_industy_return_new <- apply(Data_1971[,industry_return], 2, lag)
lag_industy_return_new <- lag_industy_return_new[-1,]


#predict with OLS coefficients
prediction_ols <- predict(model1,newx=lag_industy_return_new)
#predict with lasso coefficients
prediction_lasso <- predict(model2,s=bestlam, newx=lag_industy_return_new)

#create a new data frame to store return info
mkt_excess_new <-Data_1971$mkt_excess
return2017 <- data.frame(mkt_excess_new)
return2017$ols_excess <- 0
return2017$lasso_excess <- 0
return2017<-return2017[2:(dim(return2017)[1]),]


#calculate ols return
for (i in 1:nrow(return2017)){
  if (prediction_ols[[i]]>=0){
    return2017$ols_excess[[i]]<-return2017$mkt_excess_new[[i]]
  } else {
    return2017$ols_excess[[i]]<-return2017$mkt_excess_new[[i]]*(-1)
  }
}

#calculate ols gross exces return
for (i in 1:nrow(return2017)){
  ols_gross_return <- 1
  ols_gross_return <- ols_gross_return*(1+return2017$ols_excess[[i]])
}
print(ols_gross_return)

#calculate ols sharpe ratio
ols_sharpe <- mean(return2017$ols_excess)/sd(return2017$ols_excess)
print(ols_sharpe)


#calculate lasso return
for (i in 1:nrow(return2017)){
  if (prediction_lasso[[i]]>=0){
    return2017$lasso_excess[[i]]<-return2017$mkt_excess_new[[i]]
  } else {
    return2017$lasso_excess[[i]]<-return2017$mkt_excess_new[[i]]*(-1)
  }
}

#calculate lasso gross exces return
for (i in 1:nrow(return2017)){
  lasso_gross_return <- prod(1+return2017$lasso_excess[[i]])-1
}
print(lasso_gross_return)

#calculate lasso sharpe ratio
lasso_sharpe <- mean(return2017$lasso_excess)/sd(return2017$lasso_excess)
print(lasso_sharpe)
```


```{r}
#Plot the history of gross excess return during 1971 for both strategies and for holding market portfolio
```






